A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species

### 2016-10-11 doc/paper1-submission/paper4.tex
Update.

The flag-cleaned version of OTT contained 2,424,255 leaves.  The
\otl{} supertree was constructed from 482 phylogenetic inputs,
containing a total of 45385 leaves, of which 41029 were unique.
After flag-cleaning and exemplification by propinquity, these trees contained 40323
unique tips. We used the same cleaning flags to trim the taxonomy and
input trees when constructing \otlprop{}, so \otlprop{} contains the
same number of leaves.

@@

We performed an annotation of both the \otl{} tree and the \otlprop{}
tree as described in section \ref{sec:Annotation} to assess the
ability of our new supertree method to represent splits from input
phylogenies.  Table \ref{tab:representinputs} classifies the input
phylogeny splits according to how they relate to a summmary tree, so
that each input edge falls in one of \supportedby{}, \partialpathof{},
\terminal{}, \conflictswith{}, or \resolvedby{}. For example, the
numbers in the \conflictswith{} column indicate the number of input
splits $j$ with at least one summary edge $x$ such that the relation
``$x$ \conflictswith{} $j$'' holds.  The total number of non-terminal
input phylogeny splits considered thus 40996.

The number of displayed input splits (\supportedby{}
+ \partialpathof{}) increased from 35,518 (for \otl{}) to 39,639 (for \otlprop{});
an 11.2\% increase. When examining which splits are displayed, we find
that the \otlprop{} tree displays 4920 input splits that
are not displayed by the \otl{} tree, whereas the \otl{} tree displays
only 799 input splits that are not displayed by the \otlprop{} tree.  The number of input
splits that conflict with the summary (conflicts\_with) dropped from
2760 to 1357, a decrease of 1357, or 49.2\%.  In accordance with the
goal of not containing unnecessary polytomies, the number of input
splits that do not conflict with the summary tree, but are not
incorporated (resolved\_by) dropped from 2718 to 0.  We also find that
the number of polytomies in input phylogenies that are resolved by the
summary tree increases from 473 for \otl{} to 515 for \otlprop{}.

We also performed an annotation of the \otl{} tree and the \otlprop{}
tree to assess the degree to which these trees represent taxonomy
splits (Table \ref{tab:representtaxonomy}).  The \otlprop{} tree
conflicts with 281 more taxonomy splits than the \otl{} tree.  Since
the taxonomy is the lowest ranked input tree, this increased conflict
with the taxonomy is an expected result of incorporating more splits
from input phylogenies.

@@

Here we have described the motivation and methodology used by our new
supertree method that is currently used by the Open Tree of Life
project to build its summary tree.  Our new method represented 11\%
more input phylogeny splits with 49\% less conflict compared to the
Open Tree of Life version 4 summary tree when applied to the same inputs.
Unless the previous method \citep{HinchliffEtAl2015}, our new method
is guaranteed to incorporate input splits that do not conflict with
the summary tree. The method is implemented in the Open Source
software package propinquity.  A modified version of the treemachine
software which built the summary tree described in the
\citet{HinchliffEtAl2015} paper is used by the project to serve the
tree produced by propinquity via Open Tree of Life APIs.

>>>>

The flag-cleaned version of OTT used in the construction of both
supertrees contained 2,424,255 leaves.  The
\otl{} supertree was constructed from 482 phylogenetic inputs,
containing a total of 45385 leaves, of which 41029 were unique.
After flag-cleaning and exemplification by propinquity, these trees contained 40323
unique tips. We used the same cleaning flags to trim the taxonomy and
input trees when constructing \otlprop{}, so \otl{} and \otlprop{} contain the
same number of leaves.

@@

We performed an annotation of both the \otl{} tree and the \otlprop{}
tree as described in section \ref{sec:Annotation} to assess the
ability of our new supertree method to represent splits from input
phylogenies.  Table \ref{tab:representinputs} classifies the input
phylogeny splits according to how they relate to a summmary tree, so
that each input edge falls in one of \supportedby{}, \partialpathof{},
\terminal{}, \conflictswith{}, or \resolvedby{}. For example, the
numbers in the \conflictswith{} column indicate the number of input
splits $j$ with at least one summary edge $x$ such that the relation
``$x$ \conflictswith{} $j$'' holds.  The total number of non-terminal
input phylogeny splits considered was 40996.

The number of displayed input splits (\supportedby{}
+ \partialpathof{}) increased from 35,518 (for \otl{}) to 39,639 (for \otlprop{});
an 11\% increase. When examining which splits are displayed, we find
that the \otlprop{} tree displays 4920 input splits that
are not displayed by the \otl{} tree, whereas the \otl{} tree displays
only 799 input splits that are not displayed by the \otlprop{} tree.  The number of input
splits that conflict with the summary (\conflictswith{}) dropped from
2760 to 1357, a decrease of 1403, or 51\%.  In accordance with the
goal of not containing unnecessary polytomies, the number of input
splits that do not conflict with the summary tree, but are not
incorporated (\resolvedby{}) dropped from 2718 to 0.  We also find that
the number of polytomies in input phylogenies that are resolved by the
summary tree increases from 473 for \otl{} to 515 for \otlprop{}.

We also performed an annotation of the \otl{} tree and the \otlprop{}
tree to assess the degree to which these trees represent taxonomy
splits (Table \ref{tab:representtaxonomy}).  The \otlprop{} tree
conflicts with 281 more taxonomy splits than the \otl{} tree.  Since
the taxonomy is the lowest ranked input tree, this increased conflict
with the taxonomy is an expected result of incorporating more splits
from higher-ranked input phylogenies.

@@

Here we have described the motivation and methodology used by our new
supertree method that is currently used by the Open Tree of Life
project to build summary supertrees on the scale of millions of leaves.
Our new method represented 11\% more input phylogeny splits with 51\%
less conflict compared to the Open Tree of Life version 4 summary
tree, when applied to the same inputs. Unlike the previous method
\citep{HinchliffEtAl2015}, our new method is guaranteed to incorporate
input splits unless they conflict with the summary tree. The method is
implemented in the Open Source software package propinquity.  A
modified version of the treemachine software which built the summary
tree described in the \citet{HinchliffEtAl2015} paper is used by the
project to serve the tree produced by propinquity via Open Tree of
Life APIs.
